|

### 2024-02-06 manuscript/manuscript_2.0.Rmd
Wrapping up new manuscript version

@@

@@

In dominance and aggressive interactions, a crucial piece of
information to guide individual actions is the fighting ability of interacting 
partners, which is sometimes referred to as Resource Holding Potential 
[RHP, Parker -@parker_Assessment_1974] or simply as quality. 
Responsiveness to the quality of an opponent is central 
to communication systems in which individuals 
exhibit badges of status, where a signal typically conveys quality 
[@johnstone_Badges_1993; @rohwer_Social_1975a]. 
Such signals tend to be arbitrary, however, meaning the signals could potentially be 
produced by high- and low-quality individuals alike, opening up the 
possibility of dishonest signalling and raising the question of evolutionary stability. 
Signals can be evolutionarily stable whenever low-quality individuals 
pay a higher fitness cost for carrying the signal such that it is not 
in the interest of low-quality individuals to fake quality by producing a large badge.
Signal costs inversely proportional to quality may come about as a result of the
production costs of the signal, in which case the signal works like a 'handicap'
[@botero_Evolution_2010; @johnstone_Badges_1993;@grafen_Biological_1990;@zahavi_Mate_1975].
Throughout this manuscript, we use the word 'handicap' to refer to the differential 
costs model by Grafen [-@grafen_Biological_1990]
Alternatively, the costs can come from 
the aggressive reaction of receivers when the convention of the communication 
is broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. For example in paper 
wasps, subordinate individuals with experimentally manipulated dominant-like facial
patterns received more aggression from the dominant [@tibbetts_Socially_2004]. 
In this case, costs are socially rather than developmentally mediated and are 
triggered by the violation of the convention, although how the convention
is established in the first place is unclear.

@@

We model the evolution of signals indicating individual quality 
in the context of agonistic interactions. 
For simplicity, we consider a population of haploid 
individuals with non-overlapping generations. Individuals are born 
every generation with a level of quality ($Q_i$, where $i$ is subscript 
of the population vector of size $N$) given by a number between zero and one, 
which is drawn from a truncated normal distribution 
$N( 0.5,\sigma)$. An individual with quality $0$ has the lowest RHP, 
while an individual with quality $1$ has the highest. As they 
develop, individuals produce a phenotypic signal (badge), 
the size of which depends on their quality according to a reaction norm given by 
$B_i=1/(1+e^{\alpha_i-\beta_i Q_i})$; where $\alpha_i$ and 
$\beta_i$ are individual specific traits that determine the shape of 
the reaction norm (Fig. \ref{fig:modelStruc} A). Variation in $\alpha_i$ 
and $\beta_i$ means individuals can have either uninformative (flat) or 
informative (logistic) norms. Informative reaction norms represent a 
developmental program where the phenotype of the individual is determined by
the environmental conditions under which it grows, thus causing a 
correlation between quality and signal. The size of the signal is constrained 
to take values between zero and one. We assume different values of these traits 
are given by different alleles and are inherited from mother to offspring 
unless, with a small probability ($\mu$), mutation changes the allelic value 
of the offspring by an amount drawn from a normal distribution $TN(0,\sigma_\mu)$.

After birth, individuals go through a round of 
viability selection. The survival probability of an individual is 
given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})$, where $k_1$ and $k_2$ are parameter 
values determining the shape of the survival function. 
Importantly, if $k_2=0$ survival probability is independent of quality, 
while if $k_2>0$ lower-quality individuals pay a higher price for similar-sized 
badges, fulfilling the core assumption of the handicap principle 
[@botero_Evolution_2010; @grafen_Biological_1990; @johnstone_Badges_1993].

@@

The decision of whether to escalate a fight against an interacting 
partner can depend on the size of the partner's badge, 
with the dependency being determined by the focal's experiences acquired
through a learning process. We implement learning using 
the actor-critic approach from reinforcement learning (RL) theory [@sutton_Reinforcement_2018;@quinones_Reinforcement_2020; @leimar_Learning_2019;
@leimar_Evolution_2021].
Individuals estimate the reward (pay-off) expected from  interacting 
with partners of different badge sizes (the critic in RL terminology). 
After each interaction they update the estimate of reward proportionally 
to the difference between their current estimate and the observed 
reward (prediction error $\delta$) and to the speed of learning 
($\mathrm{A}$). Furthermore, individuals express different probabilities of 
retreating/attacking depending on the badge size of their opponent 
(the actor in RL). They update the probability of 
retreating/attacking depending on whether retreating/attacking
leads to an increase in the reward estimation. 
Thus, if a focal individual decides to escalate a fight against 
an individual with a small badge and this leads to an increase in 
the reward estimation, then the focal individual will increase the probability 
of escalating fights with individuals of small badges in the future. 
Given that badge size is a real number between 0 and 1, there are infinitely 
many badge sizes. Thus, the reward estimation, as well as the probability 
of retreating/attacking, must be generalized across different values. 
To implement generalization we use the linear function approximation 
method based on radial basis functions [@sutton_Reinforcement_2018].
Specifically, we pick $c$ feature centres, which are evenly spaced values 
along the badge size interval ([0,1]) where the updates are focused. 
For simplicity, we keep the location of these feature centres 
constant and  the same for all individuals; they are stored in 
vector $\symbf{b}$ . Each of these feature centres 
is associated with a weight for reward estimation and tendency to play 
retreat in a given interaction. The reward estimation and the tendency 
to retreat are calculated (in every interaction) as the sum of the weights 
associated with each feature centre, weighted by the response 
triggered by the feature (Fig. \ref{fig:modelStruc} B, dots 
represent the feature weights). The response of each feature centre 
diminishes as a Gaussian function with the distance between the 
feature centre and the badge size of 
the partner (Fig. \ref{fig:modelStruc} B, grey line). Formally, the 
reward estimation $\hat{R}$ when the focal individual ($i$) faces 
individual $j$ is given by,
\
\begin{equation}
\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_j-b_z|}{2\theta^2})} 
\end{equation}

@@

Individuals interact as the focal individual $n$ times and the interaction 
partner is chosen at random. Thus, the expected number of interactions for 
each individual is $2n$. 
After the interaction round , individuals in the population 
reproduce with a probability proportional to their total pay-off $w_i$, which is a sum of the 
baseline pay-off ($w_0$) and all the pay-offs obtained throughout their 
life. Thus, the combination of natural selection and genetic drift 
changes the distribution of values in $\alpha$ and $\beta$ that segregates 
in the population, effectively changing  the badges expressed and the 
communication system.

@@

```{r modelStruc,echo=FALSE,out.width='100%',fig.height=5,fig.align='center',fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions. In A, the reaction norm determines the badge size, *e.g* the size of a black band on a bird's chest. Red shows an informative reaction norm, whereas blue shows a uninformative reaction norm. B individuals perceiving the signal have a behavioral reaction norm that determines their probability of retreating (black line). The reaction norm arises from generalizing the information from the feature weights (black dots). Feature weights are values associated with a particular badge size and are updated as individuals interact with each other. The generalization of the information associated with each feature weight is represented by the grey line and its axis, which shows that the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center. The learning process moves the feature weights (black dots) up or down depending on whether this leads to an increase in the estimated reward. Panels C and D show the effect of learning on the receiver strategy. Receivers in C face signallers with uninformative reaction norms (blue line in A). The uninformative reaction norm yields a badge size of 0.5, thus the behavioural reaction norm in C is only updated around that value. D Receivers face signallers with informative reaction norms (red line in A). Accordingly, receivers develop a threashold-like reaction norm where the decision to retreat increases with the badge size of the interacting partner. Colour scale in C and D indicates the quality of the individual." ,warning=FALSE}

@@

```{r handicap,echo=FALSE,out.width='100%',fig.height=7,fig.align='center',fig.show='hold',fig.cap="Evolution of the badge size as a handicap mediated by learning. Top panels (A.1 and A.2) show the evolutionary dynamics of the sender code. In the y axis the trait value, in the x axis the generation time in thousands of generations. On the left (A.1), changes in the distribution of values for the intercept of the reaction norm ($\\alpha$); on the right (A.2) changes in the distribution of the slope ($\\beta$). Darker areas of the background correspond to values with high frequency, while orange lines in both panels show the mean of the distribution. Grey lines show the generation time corresponding to the panels below portraing the sender (B) and receiver code (C). In the bottom panels (C1-4), the learned reaction norms correspond to individuals in the population after the interaction round. Color scale indicates quality just as in Fig. \\ref{fig:modelStruc}. The black dashed line shows the average reaction norm of the population. On the first half of the simulation (B.1 and B.2) the badge size evolves to its minimum value. In the second half, the slope takes positive values and reaction norms evolve to produce an honest signal of quality. Correspondingly, individuals learn to react to badge size by increasing the probability of retreating with increasing badge sizes"  ,warning=FALSE}

@@

@@

mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
legend("topright",legend = c(expression(alpha[s])), bty = "n",cex=cexAxis)
legend("topleft",legend = c("A.1"),cex=cexAxis,bty = "n",inset = -0.06)

@@

### Badges as Handicaps

When we allowed the badge size to evolve ($\alpha$ and $\beta$ 
changed subject to natural selection and genetic drift) and the signal
worked as a handicap (*i.e.* the cost of the badge was inversely 
proportional to the quality of the individual), the sender code often evolved 
to produce an honest signal (Fig. \ref{fig:handicap}). The evolution of 
the sender code did not happen immediately after the start of the 
evolutionary process. Instead, the evolutionary dynamics of the reaction 
norm parameters (Fig. \ref{fig:handicap}, A) appeared to involve a set of 
steps. First, the intercept of the reaction norm ($\alpha$) 
describing the relationship between individual quality and badge size evolved to 
higher values, reducing the average badge size in the populations. 
This is because reaction norms segregating in the population are 
flat at the beginning of the simulation; consequently, 
the receiver code does not respond to the badge size 
and larger badges are costlier and do not trigger lower attack 
probabilities. During the first generations, the slope of the sender 
reaction norm ($\beta$)  remains close to the initial value of zero. 
At around generation 4000, the slope evolves toward positive values. 
Positive values in the slope imply that larger badges correlate 
with higher quality (Fig. \ref{fig:handicap} B). Receivers then learn to react to such 
correlation, increasing the probability of retreating when facing 
individuals with larger badges (Fig. \ref{fig:handicap} C). Hence, natural selection 
favours larger values of $\beta$, eventually leading to an evolutionary 
equilibrium in which badge size is an honest signal of quality mediated
by the learned responses of receivers (Fig. \ref{fig:handicap} B.4 and C.4).

The evolutionary trajectory portrayed in figure 
\ref{fig:handicap} A is not the only possible outcome. If the slope of the 
sender reaction norms ($\beta$) evolves, subject to genetic drift,
toward negative values before badges become handicaps, 
receivers never learn to react to the size of badges. 
Thus, badges are only costly and do not provide information about the 
quality of individuals (Fig. \ref{fig:noBadge}) . Eventually, therefore, badges 
disappear from the population. In contrast, when the badge does not 
work as a handicap but instead is cost-free, the evolutionary process 
never leads to the establishment of an honest signal. 
Instead, subject to genetic drift, the 
population either evolves towards the disappearance of the badge or to the 
maximum badge size. In either of these cases reaction norms are flat, so the 
badge does not provide any information about quality.

### The evolution of polymorphism mediated by learning

The amount of information that agents are able to collect via learning 
over their life time can strongly change the outcome of 
evolutionary dynamics. In the simulations presented in figure 
\ref{fig:handicap} and \ref{fig:noBadge}, 
individuals learned with high speed ($\mathrm{A} = 0.4$) and interacted 
repeatedly along their lifetime (2000 interactions on average). When we reduced 
the number of interactions that individuals experienced over their lifetime to 300 
on average, we saw a drastic increase in the phenotypic and genetic
variation present in the population. Populations start monomorphic 
with a value of zero for both the intercept and
the slope of the reaction norm, and mutations quickly build up a 
normal distribution around the starting 
value. Within the first 2000 generations, the unimodal distribution in 
the intercept ($\alpha$) splits into a bimodal one. 
Later in evolutionary time, one of the peaks splits further, 
so at the end of the evolutionary simulation the distribution of the intercept 
($\alpha$) in the population shows three distinct peaks. In the case of $\beta$, the 
peaks in the distribution are not so clear-cut, but the variance of the 
distribution increases over time. These changes in the 
distribution of the parameters of the sender reaction norm 
imply that individuals can generally be classified into three 
distinct types (Fig. \ref{fig:branching}). Two types express a flat reaction 
norm with extreme values for the badge size, meaning that their badge size 
is not informative of their quality, whereas the third type shows 
intermediate badge sizes determined by individual quality 
(Fig. \ref{fig:branching} B.3-4). 
In this simulation, the values of the slope in the informative group
are negative, which implies that badge size correlates negatively with quality. 
Furthermore, the receiver reaction norms developed through learning, particularly those 
of individuals of intermediate quality, respond to the signal of their 
sender type by increasing the probability of retreating from a fight with 
individuals with smaller badges (Fig. \ref{fig:branching}).

```{r branching,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of cheap signals. Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig. \\ref{fig:handicap}. The top panels (A) show changes in the distribution of values for $\\alpha$ (A.1) and $\\beta$ (A.2) along evolutionary time. In the y axis the trait value, in the x axis the generation time in thousands of generations. Panels below correspond to snapshots of the sender (B) and receiver codes (C). Generation time of the snapshots are indicated by the grey lines in the top panels. Color of the lines indicates the quality of the individual, as indicated in the color code in C.1. The dashed line represents the average reaction norm of the popúlations. The unimodal distribution with which $\\alpha$ (A.1) starts the simulation, quickly turns into a bimondal distribution, and at about 10000 generations it splits further into three modes. These three peaks correspond to three type of reaction norms in panels B.1-3."}

@@

text(x = 2000,y=-3.2,"time:")
axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
     hadj = 1.3,line = -0.5,tick = FALSE)

@@

A limited number of interactions has an effect on the amount of variation in
the evolving parameters when the signal follows the handicap
principle as well. In fig. \ref{fig:honestDiv}, we show simulations where individuals 
have on average 300 interactions in their life and the cost of the signal
is proportional to quality, following the handicap principle. 
The evolutionary process leads to a combination of reaction norm parameters where 
there is a positive correlation between the size of the badge and the quality 
of the individual. This relationship, however, is muddled by the fact that there 
are two clusters of values for the intercept ($\alpha$) and slope ($\beta$) 
of the reaction norm in the population. Thus, there are two types 
of reaction norms. One of the types has a steeper slope compared to the other. 
This effect of an increased variance in the trait distribution
is not only triggered by lower number of interactions. 
Larger variances are also found when we assume 
a lower speed of learning (data not shown). This suggest
that limits to the amount of information that individuals acquire through
learning allows the coexistence of different communication strategies within 
a population, something previously shown by @botero_Evolution_2010.

@@

mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
legend("topright",legend = c(expression(alpha[s])), bty = "n",cex=cexAxis)
legend("topleft",legend = c("A.1"),cex=cexAxis,bty = "n",inset = -0.06)

@@

@@

### The peaceful, the aggressive and the clever

>>>>

    \usepackage{geometry}

@@

    \geometry{top=1in,left=1in,bottom=1in,right=1in}

@@

In dominance and aggressive interactions, a crucial piece of
information to guide individual actions is the fighting ability of interacting 
partners, which is sometimes referred to as Resource Holding Potential 
[RHP, Parker -@parker_Assessment_1974] or simply as quality. 
Responsiveness to the quality of an opponent is central 
to communication systems in which individuals 
exhibit badges of status, where a signal typically conveys quality 
[@johnstone_Badges_1993; @rohwer_Social_1975a]. 
Such signals tend to be arbitrary, however, meaning the signals could potentially be 
produced by high- and low-quality individuals alike, opening up the 
possibility of dishonest signalling and raising the question of evolutionary stability. 
Signals can be evolutionarily stable whenever low-quality individuals 
pay a higher fitness cost for carrying the signal such that it is not 
in the interest of low-quality individuals to fake quality by producing a large badge.
Signal costs inversely proportional to quality may come about as a result of the
production costs of the signal, in which case the signal works like a 'handicap'
[@botero_Evolution_2010; @johnstone_Badges_1993;@grafen_Biological_1990;@zahavi_Mate_1975].
Throughout this manuscript, we use the word 'handicap' to refer to the differential 
costs model by Grafen [-@grafen_Biological_1990]; however, note that the 
accuracy of this terminology has been questioned [@penn_Handicap_2020]. 
Alternatively, the costs can come from the aggressive reaction of receivers 
when the convention of the communication 
is broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. For example, in paper 
wasps, subordinate individuals with experimentally manipulated dominant-like facial
patterns received more aggression from the dominant [@tibbetts_Socially_2004]. 
In this case, costs are socially rather than developmentally mediated and are 
triggered by the violation of the convention, although how the convention
is established in the first place is unclear.

@@

We model the evolution of signals indicating individual quality 
in the context of agonistic interactions. 
For simplicity, we consider a population of haploid 
individuals with non-overlapping generations of size $N=3000$. Individuals are born 
every generation with a level of quality $Q_i$ (where $i$ is subscript 
of the population vector) given by a number between zero and one, 
which is drawn from a truncated normal distribution 
$TN( 0.5,\sigma)$. An individual with quality $0$ has the lowest RHP, 
while an individual with quality $1$ has the highest. As they 
develop, individuals produce a phenotypic signal (badge), 
the size of which depends on their quality according to a logistic reaction norm given by
\begin{equation}
B_i=1/(1+e^{-\alpha_i-\beta_i Q_i}) \textrm{;}\label{eq:reacNorm}
\end{equation}

where $\alpha_i$ and $\beta_i$ are individual specific traits 
that determine the shape of the reaction norm (Fig. \ref{fig:modelStruc} 
A). Variation in $\alpha_i$ 
and $\beta_i$ means individuals can have either uninformative (flat) or 
informative (logistic) norms. Informative reaction norms represent a 
developmental program where the phenotype of the individual is determined by
the environmental conditions under which it grows, thus causing a 
correlation between quality and signal. The size of the signal is constrained 
to take values between zero and one. We assume different values of these traits 
are given by different alleles and are inherited from mother to offspring 
unless, with a small probability ($\mu$), mutation changes the allelic value 
of the offspring by an amount drawn from a normal distribution $N(0,\sigma_\mu)$.

After birth, individuals go through a round of 
viability selection. The survival probability of an individual is 
given by

\begin{equation}
s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})\textrm{,} \label{eq:viability}
\end{equation}

where $k_1$ and $k_2$ are parameter 
values determining the shape of the survival function. 
Importantly, if $k_2=0$ survival probability is independent of quality, 
while if $k_2>0$ lower-quality individuals pay a higher price for similar-sized 
badges, fulfilling the core assumption of the handicap principle 
[@botero_Evolution_2010; @grafen_Biological_1990; @johnstone_Badges_1993].

@@

The decision of whether to escalate a fight against an interacting 
partner depends on the size of the partner's badge, with the dependency 
being determined by the focal's experiences acquired
through a learning process. Individuals use previous experiences to estimate
the reward expected from encountering a partner with a particular badge size. 
However, given that badge size is a real number between 0 and 1, previous 
experiences must be generalized across the badge size range. 
To implement generalization we use the linear function approximation 
method based on radial basis functions [@sutton_Reinforcement_2018].
Specifically, we pick $c$ feature centres, which are evenly spaced values 
along the badge size interval ([0,1]) where the updates are focused. 
For simplicity, we keep the location of these feature centres 
constant and  the same for all individuals; they are stored in 
vector $\symbf{b}$. Each of these feature centres 
is associated with a weight for reward estimation 
($x_z$, $z$ is an index that goes from 1 to $c$) and tendency to play 
retreat in a given interaction ($y_z$). For each badge size ecountered the reward 
estimate and the associated tendency to retreat are calculated 
as the sum of the weights associated with each feature centre, 
weighted by the response triggered by the feature (Fig. \ref{fig:modelStruc} B, 
dots  represent the feature weights). The response of each feature centre 
diminishes as a Gaussian function with the distance between the 
feature centre and the badge size of 
the partner (Fig. \ref{fig:modelStruc} B, grey line). Formally, the 
reward estimation $\hat{R}$ when the focal individual ($i$) faces 
individual $j$ is given by,

\begin{equation}
\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_j-b_z|}{2\theta^2})} 
\end{equation}

@@

The weights vectors for reward estimation ($\symbf{x}$) and tendency 
to retreat ($\symbf{y}$) are the focus of the learning update. In the baseline
implementation of the model, these two vectors are initialized with values
of zero. For the reward estimation ($\symbf{x}$), this means that
before any interaction individuals assume they will get a payoff of 
zero. As for the behavioral policy, these initial values
translate into a $0.5$ probability of playing dove. Beyond the baseline scenario,
we explore initializing the policy vector with a) more positive values triggering 
more peaceful behaviour, b) more negative values triggering more aggressive 
behaviour, c) values that trigger the ESS of the original *hawk-dove* game and 
d) letting these values be defined by evolutionary processes 
(see Results for more details). The later was 
implemented by letting an extra locus pass on the initial value from parent to 
offspring and changed by mutation. Regardless of the initialization, all these 
values change with every interaction as a function of the reward experienced. 
Formally the change is given by,

\begin{equation}
  \Delta x_z = \Alpha \delta e^{(-\frac{|B_j-b_z|}{2\theta^2})} \textrm{.}
  \label{eq:critic}
\end{equation}

Where $\delta$ is the *prediction error*, which is the difference between the 
predicted and experienced reward, formally

\begin{equation}
  \delta = R - \hat{R} \textrm{.} \label{eq:predError}
\end{equation}

The experienced reward ($R$) is the outcome from the *hawk-dove* 
game described earlier.

Thus, the learning update depends on the speed of learning ($A$) 
and the prediction error. The exponential term assures, just as in 
the reward estimation, that the size of update is weighted by 
the distance of the partner's badge size to each feature centre.

The weights for the probability of retreat ($\symbf{z}$) are updated in a similar
fashion according to

\begin{equation}
  \Delta y_z = \Alpha \delta (1-\pi) e^{(-\frac{|B_j-b_z|}{2\theta^2})} \textrm{.}
  \label{eq:actor}
\end{equation}

where $\pi$ is the probability of choosing the action chosen by the individual. 
If an individual chooses to behave as a dove and avoid an overt conflict,

\begin{equation}
  \pi = q_{ij} = \frac{1}{1+e^{\sum_{z=1}^{c} y_z e^{-\frac{|B_i-b_z|}{2\theta^2}}}} \textrm{;}\label{eq:pi}
\end{equation}

if alternatively, the individual chooses to escalate the fight, $\pi$ corresponds 
to $1-q_{ij}$. The combination of $\pi$ and the *prediction error* ($\delta$)
assures that behavioral policy updates are carried increasing the probability of 
behaviours that predict higher reward.

This implementation corresponds to the *actor-critic* approach from reinforcement learning (RL) theory [@sutton_Reinforcement_2018;@quinones_Reinforcement_2020; @leimar_Learning_2019;@leimar_Evolution_2021]. In the *actor-critic* approach
the agent updates its behavioral policy (*actor*) and in parallel its 
reward estimation (*critic*). In our model the actor update is implemented by equations \ref{eq:actor} and \ref{eq:pi}; while the *critic* update 
is implemented by equation \ref{eq:critic}. 
Both of these updates include the *prediction error* 
(equation \ref{eq:predError}). The *prediction error* corresponds to the Rescorla-Wagner 
rule [@rescorla_Theory_1972], which is a key component of learning 
theory and widely applied in neurobiology to understand learning processes
[@holland_Minireview_2016].

Individuals interact as the focal individual $n$ times and the interaction 
partner is chosen at random. Thus, the expected number of interactions for 
each individual is $2n$. 
After the interaction round, individuals in the population 
reproduce with a probability proportional to their total pay-off $w_i$, 
which is a sum of the baseline pay-off ($w_0$) and all the pay-offs 
obtained from interactions throughout their life. Thus, the total fitness
is composed of the viability component (equation \ref{eq:viability}) and the 
total payoff collected in the interaction round. Therefore, the combination of natural 
selection and genetic drift changes the distribution of 
values in $\alpha$ and $\beta$ that segregates 
in the population, effectively changing  the badges expressed and the 
communication system.

@@

```{r modelStruc,echo=FALSE,out.width='100%',fig.height=5,fig.align='center',fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions. In A, the reaction norm  (equation \\ref{eq:reacNorm}) determines the badge size, \\textit{e.g} the size of a black band on a bird's chest. Red shows an informative reaction norm, whereas blue shows a uninformative reaction norm. B individuals perceiving the signal have a behavioral reaction norm that determines their probability of retreating (black line). The reaction norm arises from generalizing the information from the policy feature weights (black dots). Feature weights are values associated with a particular badge size and are updated as individuals interact with each other. The generalization of the information associated with each feature weight is represented by the grey line and its axis, which shows that the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center. The learning process moves the feature weights (black dots) up or down depending on whether this leads to an increase in the estimated reward. Panels C and D show the effect of learning on the receiver strategy. Receivers in C face signallers with uninformative reaction norms (blue line in A). The uninformative reaction norm yields a badge size of 0.5, thus the behavioural reaction norm in C is only updated around that value. D Receivers face signallers with informative reaction norms (red line in A). Accordingly, receivers develop a threashold-like reaction norm where the decision to retreat increases with the badge size of the interacting partner. Colour scale in C and D indicates the quality of the individual." ,warning=FALSE}

@@

```{r handicap,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap="Evolution of the badge size as a handicap mediated by learning. Top panels (A.1 and A.2) show the evolutionary dynamics of the sender code. In y axis shows the trait value, in the x axis shows the generation time in thousands of generations. On the left (A.1), changes in the distribution of values for the intercept of the reaction norm ($\\alpha$); on the right (A.2) changes in the distribution of the slope ($\\beta$). Darker areas of the background correspond to values with high frequency, while orange lines in both panels show the mean of the distribution. Grey lines show the generation time corresponding to the panels below portraing the sender (B) and receiver code (C). In the bottom panels (C1-4), the learned reaction norms correspond to individuals in the population after the interaction round. Color scale indicates quality just as in Fig. \\ref{fig:modelStruc}. The black dashed line shows the average reaction norm of the population. On the first half of the simulation (B.1 and B.2) the badge size evolves to its minimum value. In the second half, the slope takes positive values and reaction norms evolve to produce an honest signal of quality. Correspondingly, individuals learn to react to badge size by increasing the probability of retreating with increasing badge sizes."  ,warning=FALSE}

@@

evol[,meanAlpha:=meanAlpha*-1]
pop[,alpha:=alpha*-1]

@@

mtext(text = "Trait value",cex = cexAxis,line = 1.2,side = 2,las=0)
legend("topright",legend = c(expression(alpha[s])), bty = "n",cex=cexAxis)
legend("topleft",legend = c("A.1"),cex=cexAxis,bty = "n",inset = -0.06)

text(x = 2000,y=-4.2,"time:")
axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
     hadj = 1.3,line = -0.5,tick = FALSE)

@@

### Badges as Handicaps (costly signals)

When we allowed the badge size to evolve ($\alpha$ and $\beta$ 
changed subject to natural selection and genetic drift) and the signal
worked as a handicap (*i.e.* the cost of the badge was inversely 
proportional to the quality of the individual), the sender code often evolved 
to produce an honest signal (Fig. \ref{fig:handicap}). The evolution of 
the sender code did not happen immediately after the start of the 
evolutionary process. Instead, the evolutionary dynamics of the reaction 
norm parameters (Fig. \ref{fig:handicap}, A) appeared to involve a set of 
steps. First, the intercept of the logistic reaction norm ($\alpha$) 
describing the relationship between individual quality and badge size evolved to 
lower values, reducing the average badge size in the populations. 
This is because reaction norms segregating in the population are 
flat at the beginning of the simulation; consequently, 
the receiver code does not respond to the badge size 
and larger badges are costlier and do not trigger lower attack 
probabilities. During the first generations, the slope of the sender 
reaction norm ($\beta$)  remains close to the initial value of zero. 
At around generation 4000, the slope evolves toward positive values. 
Positive values in the slope imply that larger badges correlate 
with higher quality (Fig. \ref{fig:handicap} B). Receivers then learn to react to such 
correlation, increasing the probability of retreating when facing 
individuals with larger badges (Fig. \ref{fig:handicap} C). Hence, natural selection 
favours larger values of $\beta$, eventually leading to an evolutionary 
equilibrium in which badge size is an honest signal of quality mediated
by the learned responses of receivers (Fig. \ref{fig:handicap} B.4 and C.4).

The evolutionary trajectory portrayed in figure 
\ref{fig:handicap} A is not the only possible outcome. If the slope of the 
sender reaction norms ($\beta$) evolves, subject to genetic drift,
toward negative values before badges become handicaps, 
receivers never learn to react to the size of badges. 
Thus, badges are only costly and do not provide information about the 
quality of individuals (Fig. \ref{fig:noBadge}) . Eventually, therefore, badges 
disappear from the population. In contrast, when the badge does not 
work as a handicap but instead is cost-free, the evolutionary process 
never leads to the establishment of an honest signal. 
Instead, subject to genetic drift, the 
population either evolves towards the disappearance of the badge or to the 
maximum badge size. In either of these cases reaction norms are flat, so the 
badge does not provide any information about quality. This means that learning
is not sufficient to yield a communication system, but rather is requires 
the badge to have a cost proportional to the quality of the individual.

We have described here  two replicate simulations representative of the 
two outcomes found in a series of 20 replicates run with the same parameter
combination. The exact evolutionary trajectory differs in each of these replicates 
due to the stochastic effects of the dynamical system. Particularly the timing 
of the changes was different in each replicate. Nevertheless, all of the 
replicates could be classified as either evolving a functioning communication 
system, or not having the badge present in the population.

### The evolution of polymorphism mediated by learning when the signal is cost-free

The amount of information that agents are able to collect via learning 
over their life time can strongly change the outcome of 
evolutionary dynamics. In the simulations presented in figure 
\ref{fig:handicap} and \ref{fig:noBadge}, 
individuals learned with high speed ($\mathrm{A} = 0.4$) and interacted 
repeatedly along their lifetime (2000 interactions on average). When we reduced 
the number of interactions that individuals experienced over their lifetime to 300 
on average, we saw a drastic increase in the phenotypic and genetic
variation present in the population. Populations start monomorphic 
with a value of zero for both the intercept and
the slope of the reaction norm, and mutations quickly build up a 
normal distribution around the starting 
value. Within the first 2000 generations, the unimodal distribution in 
the logistic intercept ($\alpha$) splits into a bimodal one. 
Later in evolutionary time, one of the peaks splits further, 
so at the end of the evolutionary simulation the distribution of the 
logistic intercept ($\alpha$) in the population shows three distinct peaks. 
In the case of $\beta$, the peaks in the distribution are not so clear-cut, 
but the variance of the distribution increases over time. 
These changes in the distribution of the parameters of the sender reaction norm 
imply that individuals can generally be classified into three 
distinct types (Fig. \ref{fig:branching}). Two types express a flat reaction 
norm with extreme values for the badge size, meaning that their badge size 
is not informative of their quality, whereas the third type shows 
intermediate badge sizes determined by individual quality 
(Fig. \ref{fig:branching} B.3-4).

In this simulation, the values of the slope in the informative group
are negative, which implies that badge size correlates negatively with quality. 
The receiver reaction norms developed through learning, particularly those 
of individuals of intermediate quality, respond to the signal of their 
sender type by increasing the probability of retreating from a fight with 
individuals with smaller badges (Fig. \ref{fig:branching}). However, 
given the absence of costs, the sign of the slope is arbitrary. In our replicate 
simulations we found examples where the third type had positive 
and negative slopes.

```{r branching,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of cheap signals. Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig. \\ref{fig:handicap}. The top panels (A) show changes in the distribution of values for $\\alpha$ (A.1) and $\\beta$ (A.2) along evolutionary time. The y axis shows the trait value, and  the x axis shows the generation time in thousands of generations. Panels below correspond to snapshots of the sender (B) and receiver codes (C). Generation time of the snapshots are indicated by the grey lines in the top panels. Color of the lines indicates the quality of the individual, as indicated in the color code in C.1. The dashed line represents the average reaction norm of the populations. The unimodal distribution with which $\\alpha$ (A.1) starts the simulation, quickly turns into a bimondal distribution, and at about 10000 generations it splits further into three modes. These three peaks correspond to three type of reaction norms in panels B.1-3."}

@@

text(x = 2000,y=-3.3,"time:")
axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
     hadj = 1.3,line = -0.5,tick = FALSE)

@@

A limited number of interactions has an effect on the amount of variation in
the evolving parameters when the signal follows the handicap
principle as well. In fig. \ref{fig:honestDiv}, we show simulations where individuals 
have on average 300 interactions in their life and the cost of the signal
is proportional to quality, following the handicap principle. 
The evolutionary process leads to a combination of reaction norm parameters where 
there is a positive correlation between the size of the badge and the quality 
of the individual. This relationship, however, is muddled by the fact that there 
are two clusters of values for the logistic intercept 
($\alpha$) and slope ($\beta$) of the reaction norm in the population. 
Thus, there are two types of reaction norms. 
One of the types has a steeper slope compared to the other. 
This effect of an increased variance in the trait distribution
is not only triggered by lower number of interactions. 
Larger variances are also found when we assume 
a lower speed of learning (data not shown). This suggest
that limits to the amount of information that individuals acquire through
learning allows the coexistence of different communication strategies within 
a population, something previously shown by @botero_Evolution_2010.

@@

mtext(text = "Trait value",cex = cexAxis,line = 1.2,side = 2,las=0)
legend("topright",legend = c(expression(alpha[s])), bty = "n",cex=cexAxis)
legend("topleft",legend = c("A.1"),cex=cexAxis,bty = "n",inset = -0.06)

@@

text(x = 2000,y=-1.2,"time:")
axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
     hadj = 1.3,line = -0.5,tick = FALSE)

@@

### The peaceful, the aggressive and the clever, when signals are cost-free
